---
title: "Untitled"
author: "Jonas Björnerstedt"
date: '`r Sys.Date()`'
output:
  slidy_presentation: default
  powerpoint_presentation: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(knitr)
library(shiny)
library(kableExtra)
library(flextable)

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(dplyr.summarise.inform = FALSE)
options(knitr.kable.NA = '') # Suppress printing of NA in kable
# Standard theme for document 
theme_set(theme_bw()) 

set_flextable_defaults(
align = "center",
digits = 2
# theme_fun = theme_zebra
  )

doShiny = TRUE

source("code/atmp.R")
# source("create_settings.R")

# Läs in från Excel
# example_simple = open_indata("Simple.xlsx")
exampleA = open_indata("code/Example_A.xlsx")
exampleB = open_indata("code/Example_B.xlsx")
# exampleA2 = open_indata("Example_A2.xlsx")
```

## A simple example

```{r}
exampleA$state_table %>% 
  # select( -p_death) %>% 
  filter(treatment == "ATMP") %>% 
  with_titles( exampleA$state_description) %>% 
  flextable_output() %>% 
  colformat_double(j = c(2), digits = 0) 
```


```{r}
library('diagram')

Numgenerations <- 4
DiffMat <- matrix(data = 0, nrow = Numgenerations, ncol = Numgenerations)
AA <- as.data.frame(DiffMat)
AA[[4,1]] <- ".01  "
AA[[4,2]] <- ".02"
AA[[4,3]] <- ".02"
#
AA[[1,1]] <- "0.95"
AA[[2,1]] <- "0.04"
AA[[3,2]] <- "0.98"
# AA[[4,3]] <- "0.98 + 0.02"
AA[[4,3]] <- "1.00"
#
name <- c("H","Pr. 1","Pr. 2", "D")
#
plotmat(A = AA, pos = c(3,1), name = name, lwd = 2, self.lwd = 2, arr.len = 0.2, relsize = .7, arr.width = 0.1, my = 0.29, box.size = 0.09, arr.type = "triangle", dtext = 0.1)
```

## Transition matrix

```{r}
  print_transitions(exampleA, treatment_name = "ATMP")
```

## Two treatments, one with certain progression

```{r}
exampleA$state_table %>% 
  with_titles( exampleA$state_description) %>% 
  flextable_output() %>% 
  colformat_double(j = c(2), digits = 0) 
```


## Payment plans

Two simple __Payment plans__

* _For ATMP tr._ - total payment of 15 spread over 10 periods

* _For comparator tr._ - continuous payment of 0.1 in all periods

```{r, message=FALSE}
exampleA$payment_table %>% 
    with_titles( exampleA$payment_description) %>% 
  flextable_output() %>% 
  colformat_double(j = c(4), digits = 0) 
```

## Structure of Payment plans

* Available specifications with default values

```{r, message=FALSE}
exampleA$payment_description %>% select(-name, -min, -max) %>% flextable_output() 
```

## Global variables

```{r}
exampleA$global_table %>% slice(1:3)  %>% flextable_output()
```

# Examples

## __Example A__ 

A basic example with two treatments, each with a payment plan

## __Example B__ 

* Similar to A, but with payment for comparator treatment if ATMP fails (by entering state 2). 

* Also includes treatment ATMP certain with zero probability of progression.

## __Example C__ 

* Comparing immediate introduction of ATMP treatment with waiting one period for a lower expected price. 

* The example also has discount rates for the firm and the HA.

## __Example D__ 

* Similar to A but with a more complicated payment plan. 
* The payment is all in the initial period, but if the treatment fails within 10 periods, there is a refund of 50%. 
* As the Agg. fail limit is set to 30%, the refund will be on all payments if 30% of the treated have reached state 2 within 10 periods. 

* As can be seen in the cost plot on the _Analysis_ page, the refund comes in period 8.

## __Example E__ 

* The ATMP treatment has a slightly higher quality of life (0.8) than the comparison alternative (0.7). 

* There is no progression for either treatment, but a probability of ATMP ceasing to work as well as a probability of dying.

## __Example F__ 

* A more complicated comparison of payment plans together with different probabilities of the ATMP treatment ceasing to work.