---
title: "Simulation of ATMP payment plans"
always_allow_html: true
output:
  md_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(knitr)
library(shiny)
library(kableExtra)
library(flextable)

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(dplyr.summarise.inform = FALSE)
options(knitr.kable.NA = '') # Suppress printing of NA in kable
# Standard theme for document 
theme_set(theme_bw()) 

set_flextable_defaults(
align = "center",
digits = 2
# theme_fun = theme_zebra
  )

doShiny = TRUE

source("atmp.R")
# source("create_settings.R")

# Läs in från Excel
# example_simple = open_indata("Simple.xlsx")
exampleA = open_indata("Example_A.xlsx")
exampleB = open_indata("Example_B.xlsx")
# exampleA2 = open_indata("Example_A2.xlsx")
```

# Introduction

The ATMP package is an R package to analyse the long term health effects and costs of ATMP treatments. The package focuses on the analysis of different payment plans that condition on outcomes over time. Modelling payments over a long time horizon, the package allows payment plans that specify payments or reimbursements that can depend on individual or aggregate outcomes. 

The current version simulates the expected health and cost effects given the parameters used. Modelling the variability outcomes in terms of individual outcomes, or uncertainty in the parameter estimates increases the level of complexity of the interface. Given the uncertainties in long term estimation, we have chosen to focus on allowing variation in parameter estimates rather than specifying a distribution for the uncertain parameters.

Two guiding principles in the design of the package are flexibility and ease of use. Flexibility is crucial as both ATMP treatments and payment models can be diverse, and the modelling of new treatments cannot be completely foreseen at the present time. The package allows for simple analysis for users less experienced in R programming as well as customisation of analysis for users that are more familiar with R and Rmarkdown. 

# Structure of a simulation

A simulation is based on a combination of _treatments_ and _payment plans_, combined with a set of _global parameters_ such as discount factors and the time horizon of the analysis. 

_Treatments_ specify Markov health models, with states, transitions between states and the quality of life of each state. The ATMP package restricts the structure of the health model in order for it to be easier to use. In this first version transitions are all in one direction - the health state never improves. Although the underlying functionality is general allowing arbitrary transitions between health states - such generality is more difficult to specify in a simple way for users.   

_Payment plans_ specify how much is payed when under what conditions. A treatment can include several payments each over several time periods. Payment plan payments can begin and end depending on health states. For example if a treatment stops working, payment can end and the payment for an alternative treatment commences.  

Payment plans and treatments are associated by a _Arm_: the active or the control arm.  
  
  - Many payment plans can be associated with an arm, for different payments of different treatments in the treatment process
  
  - Many treatments can be included in an arm, to analyse the health outcome under varying treatment assumptions 

## Model specification

Inputs to the simulation are set in an Excel file. Each health case has an Excel file specifying health and payment parameters for the models used. Although all parameters are set in the Excel sheet, one can also allow modification of parameters interactively using the ATMP app. Although many parameters can be set in the Excel tables, only the parameters of interest have to be included in the tables. With sensible default values for parameters, those that are not central do not have to be included. 

# A simple example

Below is a minimal treatment table for the  _ATMP_ treatment, with 4 states. The probabilities _Pr. prog_ are the probabilities of progressing to the next state. The probability of moving from state 1 to 2 is thus 0.05. Each state is associated with a Quality of Life _QoL_, and a _Payment plan_. In State 1, the payment is for the ATMP treatment. In State 2 we have assumed that the ATMP treatment has stopped working, in which case the comparator treatment is used. In the final state 4, there are no payments and QoL is assumed to be zero. For simplicity, we assume in this example that the probability of dying from other reasons is zero. 

```{r}
exampleA$state_table %>% 
  select( -p_death) %>% 
  filter(treatment == "ATMP") %>% 
  with_titles( exampleA$state_description) %>% 
  flextable_output() %>% 
  colformat_double(j = c(2), digits = 0) 
```

For the ATMP treatment, the transition from full health (state 1) to progression (state 2) is random, with a probability of 0.05 of moving to state 2.  Progression once started is deterministic, with probability 1 of moving to the next state, eventually ending in death (state 4). 

Below is the treatment table for two treatments: _ATMP_ and _Comparison_. We have now included a probability of death _Pr.death_. The only difference between the treatments is in state 1. For the _Comparison_ treatment the comparator drug is used and the probability of transitioning to state 2 is 1 as indicated in the column _Pr.prog_.  
 

```{r}
exampleA$state_table %>% 
  with_titles( exampleA$state_description) %>% 
  flextable_output() %>% 
  colformat_double(j = c(2), digits = 0) 
```

The transition matrix is given by

```{r}
  print_transitions(exampleA, treatment_name = "ATMP")
```

Treatment tables can include more detailed information. At this time the following variables can be specified in a treatment table. If not specified, the default values in the table below are used (given by the _value_ column). For example the random state does not have to be the first state - we can have delayed treatment where there is progression prior to treatment. A restriction in the current model is that only one random state is allowed (i.e. with _Pr. prog_ between zero and one).

```{r }
exampleA$state_description %>% select(-name, -min, -max) %>% flextable_output()
```

Probabilities of progression and dying are based on the assumption that these are independent.  

## Quality of life

To each state, a measure of the quality of life (QoL) is associated. As specified both treatments have the same linear decrease in QoL from 1 (full health) to 0 (death) over the 4 states.

```{r, eval=FALSE}
plot_QoL(exampleA)
```


# Payment plans

Payment plans are specified in a table in the Excel input sheet. A simple example from above, we let each treatment have a corresponding payment plan.

```{r, message=FALSE}
exampleA$payment_table %>% 
    with_titles( exampleA$payment_description) %>% 
  flextable_output() %>% 
  colformat_double(j = c(4), digits = 0) 
```

The total _ATMP payment_ is 10 with a payment scheme length of 10 years. This implies that the payment is spread evenly over the ten years, with a payment of 1 each year. The payment scheme also specifies that payments cease if the next state (here state 2) is reached before ten years have passed,  

For the _Comparison_ alternative we assume continuous purchases as long as the patient is alive. Payments are made as long as state 4 (death) is not reached. Payments of 0.5 have been specified as continuous payments, implying that this is the yearly payment. The payment scheme length has no effect in this case, here set to 0.

The payment plans for _ATMP payment_ and _Comparison_ over a 20 year time horizon are shown below.

```{r}
plot_payment_plans(exampleA)

```


## Structure of Payment plans

A full list of specifications is given in the table below. 

```{r, message=FALSE}
exampleA$payment_description %>% select(-name, -min, -max) %>% flextable_output() 
```

Payment plans specify:

1. Payments over time such as payment length, share of payments as initial payment.

2. Conditions for payments ending can be specified in different ways

  - Individual payments cease if a health state is reach, such as progression starting and/or at death
  
  - Payments for all treated can cease if the aggregate failure of the treatment is above a specified threshold
  
  - Cessation of payments can also include a refund, specified as a share of payments.
  
3. The state where payments start can be specified

  - Ex. If ATMP stops working, the cost of an alternative treatment can be included. 
  
4. If Tot.payment is specified, continuous payment is not used. The payment is distributed over the Periods. An initial payment can be set, specifed as a percentage of the total payment.

## Global variables

In addition to specifying treatments and payment plans, _global variables_ are also specified in a table. All parameters common to all treatments and payment plans in the case are specified in the global table. The time horizon of analysis, here 20 years, and discount rate are such common parameters.

```{r}
exampleA$global_table %>% flextable_output()
```

# Analysis

Analysis of health effects and costs can be done either by simply processing the input data specified in the Excel file, or interactively in the ATMP app.

The results of the analysis can be 

1) Shown on screen

2) Generated as a report in a chosen format: HTML/Word/Pdf/Powerpoint. 

3) As Excel tables, for further analysis

Interactive analysis allows modification of all input specified in Excel sheet. This allows interactive change of important parameters such as payment scheme length or ATMP hazard. The modified model can be saved to an excel file.
  
As even simple health models require a lot of indata, interactivity is based on setting all values of interest in an Excel sheet. A starting point for a model can be one of the example models provided with the app. The Excel sheet specifies all indata, with the user only modifying selected variables. The choice of variables for interactive analysis will depend on the specific model at hand. 

## Using the ATMP app

When the ATMP app starts, you are med with a screen, with six tabs at the top:

1.  *Overview* - The current tab
2.  *Model* - Select health model
3.  *Input* - See/modify the input values of the model
4.  *Analysis* - Analysis of the health model
5.  *Report* - Download a report of the analysis
6. *Help* - This text.

### The Model tab

The starting point of the analysis is a _health model_; tables of input values specified in an Excel sheet. Selecting the _Model_ tab displays a page where you select the health model to analyse. You can open an example model by selecting one of the models in the list box, or open an Excel file from your computer.

Once loaded, the page will display figures of quality of life and payment plans of the health model.

### The Input tab

Once loaded, values can be modified under the _Input_ tab. Simply double click on a value to change it, clicking somewhere else once you have entered a new value. The model tables displayed on the _Input_ page can be downloaded as an Excel file to your computer using the _Download_ button at the bottom of the _Input_ page.

### The Analysis tab

Results of the analysis are displayed in the _Analysis_ tab. Some parameters will also affect the plots on the _Model_ page.

### The Report tab

You can download a report containing the tables and figures displayed on the Model and Analysis pages, in the desired file format.

## Example models

The ATMP app has several example models that can be loaded

__Example A__ A basic example with two treatments, each with a payment plan

__Example B__ Similar to A, but with payment for comparator treatment if ATMP fails (by entering state 2). Also includes treatment ATMP certain with zero probability of progression.

__Example C__ Comparing immediate introduction of ATMP treatment with waiting one period for a lower expected price. The example also has discount rates for the firm and the HA.

__Example D__ Similar to A but with a more complicated payment plan. The payment is all in the initial period, but if the treatment fails within 10 periods, there is a refund of 50%. As the Agg. fail limit is set to 30%, the refund will be on all payments if 30% of the treated have reached state 2 within 10 periods. As can be seen in the cost plot on the _Analysis_ page, the refund comes in period 8.

__Example E__ The ATMP treatment has a slightly higher quality of life (0.8) than the comparison alternative (0.7). There is no progression for either treatment, but a probability of ATMP ceasing to work as well as a probability of dying.

__Example F__ A more complicated comparison of payment plans together with different probabilities of the ATMP treatment ceasing to work.

These examples can be used as a starting point in creating your own model. 

## ATMP package function interface

To read the data from the Excel indata file, the function `open_indata()` is used. It takes the filename as an argument and returns a structure with the tables described above. Here this structure is put in the variable `indata`.

```{r, eval=FALSE, echo=TRUE}
indata = open_indata("models.xlsx")
```

The main function `analyse_treatments()` generates a data.frame of QALYs for each treatment and Costs for each payment scheme associated with the treatment. The tables shown on the _Analysis_ page can be generated by the following function calls

```{r, eval=FALSE, echo=TRUE}
analyse_treatments(indata) 

analyse_treatments(indata, show_details = TRUE)
```

The results of the analysis is returned in a dataframe, to be displayed as a table or for further analysis. To study the treatments over time the corresponding option is used

```{r, eval=FALSE, echo=TRUE}
df = analyse_treatments(indata, over_time = TRUE)
```

The `df` object returned by the function is a data frame that can be analysed using standard tools in R.

