---
title: "New structure"
author: "Jonas Björnerstedt"
date: "`r Sys.Date()`"
output: html_document
---

# Översikt

Model:

* plot_QoL(indata)

* plot_payment_plans(indata)

# Analysis

* contract_analysis(vals, show_details = TRUE) - analyse_treatments

* plot_costs(indata)

* plot_QALY(indata)

Funktioner:

* check_indata(indata)

Tutorial:

* plot_treatment_paths(indata, treatment_name, reps = 8, T = 20) 

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(knitr)
library(shiny)
library(kableExtra)
library(flextable)

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(dplyr.summarise.inform = FALSE)
options(knitr.kable.NA = '') # Suppress printing of NA in kable
# Standard theme for document 
theme_set(theme_bw()) 

set_flextable_defaults(
align = "center",
digits = 2
# theme_fun = theme_zebra
  )

doShiny = TRUE

source("atmp.R")
source("atmp_new.R")

# Läs in från Excel
# source("create_settings.R")
example_simple = open_indata("simple.xlsx")
exampleA = open_indata("Example_A.xlsx")
exampleB = open_indata("Example_B.xlsx")
exampleA2 = open_indata("Example_A2.xlsx")
exampleC = open_indata("Example_C.xlsx")

```

# Sannolikheter

# State table version

```{r}
state_tableA = exampleA$state_table
exampleA$state_table %>% flextable_output()
```

* Omit QoL and it is linear

* Can specify QoL at transition instead of NA. Can also specify QoL for each state.

Not done

* If state is omitted, it is set to the next value. Useful for example 3 and for omitting state completely.

* Omit p_death and it is set with real hazard

* Lägg till arm kolumn om det är fler än två treatments

```{r, eval=TRUE}
exampleA$payment_table %>% 
  flextable_output()   
```



## Create full table

Fill state table and create item variable - could be useful in payments that continue over states

```{r}
state_table = create_state_table(exampleA$state_table) %>% 
  filter(treatment == "ATMP")
state_table  %>% flextable_output()
```

## Example B

```{r}

exampleB$state_table %>% flextable_output()
```

## Example C

```{r}
exampleC$state_table %>% flextable_output()
```

# Calculations

```{r}
P = transition_matrix(state_table) 
P
```

```{r}
get_QoL(state_table) 

```

```{r}
expected_markov( P, 20)

```

For continuous payments, we can calculate costs by simply creating a cost vector - payment per state

For multi-period payments with certain starting time we can do the same as we know the time and the share entering the ATMP state. 

# Payment plans

```{r}
contract_description = read_csv("
name, title, value,               description
payment, Payment plan, ,            Payment name
tot_payment,  Tot. payment, 0,     Total payment or yearly paymentif payment is continuous
cont_payment, Cont. payment, 0, Payment each year that the patient is alive (traditional) 
cost_trend, Cost Trend,      0, Cost trend over time - for continuous payment
contract_length, Periods,   20, Total length of contract
initial_payment, Initial,   0,    Share of payment in initial period
refund, Refund,             0, Share of payments refunded upon failure (progression)
aggregate_failure, Agg. Fail, 0, Threshold share of population for aggregate failure
")

# treatment_table = read_csv("
# treatment, p_HU, health_states
# ATMP,              0.04, 4
# Comparison,        1, 4
# ") 

payment_table = exampleA$payment_table

global_table = read_csv("
name, value
discount, 0.0
firm_discount, 0.0
time_horizon, 20
threshold, 1
active_plan, 1
control_plan, 0
"
) 
```


Single contract as list with defaults

```{r}
con_def = named_list(contract_description, "name", "value")
globals = named_list(global_table, "name", "value")

```



```{r}
state_table %>% 
  left_join(payment_table) %>% 
  mutate_if(is.numeric, list(~replace_na(., 0))) %>% 
  mutate(payment = replace_na(payment, "Death")) 


```


```{r}
create_payment_plans(exampleA$state_table %>% filter(treatment =="ATMP"), exampleA)

```

# Results

does not summarise
```{r}
analyse_treatments(exampleA)  %>% flextable_output()

```

```{r}
source("atmp_new.R")
analyse_treatments(exampleA, show_details = TRUE) %>% flextable_output()
```
```{r}
contract_analysis(exampleA, show_details = TRUE)  %>% flextable_output()
```


